---
title: "Comparison Analysis for Sorting Algorithms"
author: "Samyak Ahuja"
output: pdf_document
---

```{r include=FALSE}
library(ggplot2)
library(ggpmisc)
library(reshape2)
```

# Overview

Sorting Algorithms chosen for analysis are : 

+ Insertion Sort
+ Merge Sort
+ Quick Sort


# Helper Functions

Helper functions are used for two purposes:

|Data Processing|Plotting|
|---|---|
|Data Generator|Individual Plotter|
|Comparison Finder|Combined Plotter|

## Data Generator and Comparison Finder

### Data Generator

#### About

Objective : To formulate a dataset that is same for all the Sorting functions

Input : 

+ n which is the maximum number of elements in the set
+ sep which is the separator by which the number of elements in the set are increased

Output : Dataset on which Sorting is done. The format of the dataset is explained below.

#### Dataset Structure

Dataset is a list with the following elements by row.

| 1 | 2 | ... | 10 | 
|---|---|---|---|
| X-1,1 | X-1,2 | ... | X-1,10 |
| X-2,1 | X-2,2 | ... | X-2,10 |
| ... | ... | ... | ... |
| X-n,1 | X-n,2 | ... | X-n,10 |
            
X-i,j is an array with a number Xi from 0 to 100

[X1, X2, ..., Xi]


```{r}

dataSetGenerator <- function(n = 1000, sep = 10){
    ele <- seq(from = 0, to = n, by = sep)
    ele <- ele[-1]
    data <- list()
    for(j in ele){
      iterator <- j / sep
      repeated <- list()
      for(i in 1:10){
        repeated <- c(repeated, list(sample(x = 1:100, size = j, replace = TRUE)))
      }
      data <- c(data, repeated)
    }
    return (data)
}

dataSet <- dataSetGenerator()

```


### Comparison Finder

#### About

Objective : To output the average number of comparisons used for each row in the dataset given the sorting algorithm

Input : 

+ func which is the sorting function to use on the dataset
+ n which is the maximum number of elements in the set
+ sep which is the separator by which the number of elements in the set are increased

Output : a Data-Frame (Matrix) that has two Columns :

+ ele which is the number of elements in an array given for Sorting.
+ timeElapsed which is the **average** number of comparisons used for that sorting algorithm

&nbsp;

```{r}
comp_find <- function(func, n = 1000, sep = 10){
  ele <- seq(from = 0, to = n, by = sep)
  ele <- ele[-1]
  timeElapsed <- c()
  for(j in ele){
    op <- 0
    iterator <- j / sep
    for(i in 1:10){
         op = op + func(dataSet[[iterator + i]])$operations
    }
    #taking average over 10 examples of same size
    op = op / 10
    timeElapsed <- c(timeElapsed, op)
  }
  return (data.frame(ele,timeElapsed))
}
```

## Plotting

### Individual Plotter

plotter function creates a Comparisons vs Elements plot for each sorting algorithm separately.

The Fitting is done using a polynomial curve of degree 2.

&nbsp;

```{r}
plotter <- function(df, df_title){
  ggplot(df, aes(ele, timeElapsed, color = timeElapsed)) +
    geom_point(shape = 16, size = 5, show.legend = FALSE, alpha = 0.6) + 
    stat_smooth(method="lm", formula=y~poly(x,2), rm = FALSE) + 
    theme_minimal() + 
    labs(subtitle = "Comparisons vs Size",
       y = "Number of Comparisons (Averaged)",
       x = "Number of Elements",
       title = df_title) +
    scale_color_gradient(low = "#32aeff", high = "#f2aeff") +
    stat_poly_eq(parse=T, aes(label = ..eq.label..), formula=y~poly(x,2))
}
```

### Combined Plotter

The comb_plotter function creates a combined Comparisons vs Elements plot for all the sorting algorithms.

The Fitting is done using a polynomial curve of degree 2.

&nbsp;

```{r}
comb_plotter <- function(df, df_title){
  ggplot(df, aes(ele, value, col = variable)) + 
  geom_point(shape = 16, size = 2, alpha = 0.6) + 
  stat_smooth(method="lm", formula=y~poly(x,2)) + 
  theme_minimal() +
  labs(subtitle = "Comparisons vs Size",
       y = "Number of Comparisons (Averaged)",
       x = "Number of Elements",
       title = df_title) +
  stat_poly_eq(parse=T, aes(label = ..eq.label..), formula=y~poly(x,2)) 
}
```


# Sorting Function - Implementation

## Insertion Sort

### Sorting Algorithm
```{r}
insertionSort <- function(vec){
  n <- length(vec)
  comparisons <- 0
  for(i in 2:n){
    key <- vec[i]
    pos <- i - 1
    while(pos > 0 && vec[pos] > key){
      vec[pos + 1] = vec[pos]
      pos = pos - 1
      comparisons <- comparisons + 1
    }
    vec[pos + 1] <- key
    comparisons <- comparisons + 1
  }
  return (list("vec" = vec, "operations" = comparisons))
}
```

### Proof of concept
```{r}
insertionSort(c(12,-22,13,2,-33,2))
```


## Merge Sort

### Sorting Algorithm

```{r}
mergeSort <- function(vec){
  
  #merges the two sorted halves left and right to get a sorted list
  mergeTwo <- function(left,right){
    comparisons <- 1
    res <- c()
    while(length(left) > 0 && length(right) > 0){
      comparisons <- comparisons + 1
      if(left[1] <= right[1]){
        res <- c(res,left[1])
        left <- left[-1]  
      }else{
        res <- c(res,right[1])
        right <- right[-1]
      }
    }
    if(length(left) > 0){
      res <- c(res,left)
    }
    if(length(right) > 0){
      res <- c(res,right)
    }
    return (list("vec" = res, "operations" = comparisons))
  }
  
  comparisons <- 0
  n <- length(vec)
  
  if(n <= 1) return (list("vec" = vec, "operations" = comparisons))
  else{
    middle <- length(vec) %/% 2 #integer division
    
    left_list <- mergeSort(vec[1:middle])
    right_list <- mergeSort(vec[(middle + 1):n])
    left <- left_list$vec
    right <- right_list$vec
    
    res <- mergeTwo(left,right)
    comparisons <- left_list$operations + right_list$operations + res$operations
    return (list("vec" = res$vec, "operations" = comparisons))
  }
}
```

### Proof of Concept
```{r}
mergeSort(c(12,-22,13,2,-33,2))
```


## Quick Sort

### Sorting Algorithm

```{r}
quickSort <- function(vec, low = 1, high = length(vec)){
  
  #partitioning vector around pivot(chosen as last element of the vector)
  partition <- function(vec, low, high){
    i = low
    comparisons <- 0
    pivot = vec[high]
    
    for(j in low:(high - 1)){
      comparisons <- comparisons + 1 
      if(vec[j] <= pivot){
        #swap v[i] and v[j]
        temp = vec[i]
        vec[i] = vec[j]
        vec[j] = temp
        i = i + 1
      }
    }
    #swap pivot with v[i]
    temp = vec[i]
    vec[i] = vec[high]
    vec[high] = temp
    
    return (list("vec" = vec, "operations" = comparisons, "pi" = i))
  }
  
  comparisons <- 0
  if(low < high){
    pi_list = partition(vec, low, high)
    vec <- pi_list$vec
    pi <- pi_list$pi
    
    left_list <- quickSort(vec, low, pi - 1)
    vec <- left_list$vec
    
    right_list <- quickSort(vec, pi + 1, high)
    vec <- right_list$vec
    
    comparisons <- left_list$operations + right_list$operations + pi_list$operations
    return (list("vec" = vec, "operations" = comparisons))
  }else{
    return (list("vec" = vec, "operations" = comparisons))
  }
}
```

### Proof of Concept
```{r}
quickSort(c(12,-22,13,2,-33,2))
```


# Sorting Algorithms - Plots 

## Individual Plots

### Insertion Sort
```{r}
isdf_small <- comp_find(insertionSort)
plotter(isdf_small, "Insertion Sort")
```

### Merge Sort
```{r}
msdf_small <- comp_find(mergeSort)
plotter(msdf_small, "Merge Sort")
```

### Quick Sort
```{r}
qsdf_small <- comp_find(quickSort)
plotter(qsdf_small, "Quick Sort")
```


## Combined Plots

```{r}
df_small <- data.frame(ele = msdf_small[[1]],
                 insertionSort = isdf_small[[2]], 
                 mergeSort = msdf_small[[2]], 
                 quickSort = qsdf_small[[2]])
df_small
df_small <- melt(df_small, id.vars = "ele")
comb_plotter(df_small, "Combined Scatter Plot")
```
